---
title: "The Future of Global Fishery Assessments"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2: default
  bookdown::pdf_document2: default
abstract: Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals. Numerous high-profile publications have produced estimates of the global status of fisheries, but limited quantity and quality of data, conflicting results, and methodological concerns mean that our understanding of global fisheries is still highly uncertain, especially in critical areas such as Southeast Asia and Africa. We present an evaluation of the past and future of global fishery assessments. Global assessments have typically involved aggregating best-available stock assessments where possible with various data-limited assessments for regions lacking formal evaluations of stock status. The data-limited methods used have ranged from simple catch heuristics to super-ensemble modeling. Despite attempts at model improvement, there have been discrepancies both among global fishery assessments and between the data-limited components of these methods and more data-rich models,demonstrating that our understanding of global data-limited fisheries remains limited. How then can we most efficiently move forward? We utilize a flexible assessment model together with the RAM Legacy Stock Assessment Database not strictly as a stock assessment tool, but to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion and increasingly available but historically underutilized data such as trawl footprints. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world. We close by discussing priority areas for future data collection. Our results make the case that improving our understanding of global fisheries depends not on better models, but on better data.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(here)
extrafont::loadfonts()

results_name <- "v0.5"

results_dir <- here("results",results_name)

load(file = file.path(results_dir, "paper_plots.Rdata"))

pub_theme <- theme_ipsum(base_size = 12, 
                         axis_text_size = 12)

theme_set(pub_theme)

```


Target journal: Nature Sustainability 

https://www.nature.com/natsustain/about/content

3500 words main text 
 
50 references (main text)

3000 words (max) methods

# Introduction (500 words)

*Insert mandatory opening sentence about the importance of global fisheries*. On-the ground fisheries management decisions are ideally based on assessment models, designed to provide robust estimates of individual stock status that can inform effective management policies. However, data and resource limitations prohibit our ability to create bespoke tactical assessments for every fishery resource around the globe. As such, we also depend on widely applicable methods intended to provide a general sense of the status of fisheries around the globe. While not suitable for management of individual stocks, these broad strategic models play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals. 

The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely followed estimate of global stock status. As of the most recent report, the FAO estimates that 59.9% of marine fish stocks are maximally sustainably fishery, 7% are underfished ,and 33.1% are overfished [@fao2018]. While foundational, the SOFIA reports depend on a complex process of aggregating results from tactical models, expert judgement, etc. **make clear argument why other/new global assessment methods are needed**. 

Several publications have explored alternative methods for estimate global stock status, including Costello et al. 2016 [@costello2016a]  and Rosenberg et al. 2018 [@rosenberg2018], and Pauly 2007 [@pauly2007]. Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses. However, comparing estimates from each of these methods against best-available knowledge of stock status from the RAM Legacy Stock Assessment Database [@ricard2012], as well as broader intuition. reveals some systemic problems with these assessments. 

We have stock assessment data from RAM from the NE Pacific,the NE Atlantic and the Mediterranean, and of these the NE Pacific is in the best state (mean RAM B/B~MSY~ =1.8), NE Atlantic (mean RAM B/B~MSY~ =1.3) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but based on surveys and expert opinion we believe it should be worse than Northeast Pacific. The Costello et al. 2016 [@costello2016a] model shows the correct general pattern, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison. The Rosenberg super-ensemble method[@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)) *probably want to make this a figure. Also, Ray, do you remember if these estimates are unassessed only, or unassessed + RAM?*

```{r status-tab}

readr::read_csv(here("data","status-table.csv")) %>% 
  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
               booktabs = TRUE) %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```


Despite numerous advances in statistical methods, we are currently left with strong questions about our understanding of the state of global fisheries, particularly in the regions of the world lacking formal stock assessments. We build off of these prior works to demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection effort. 


# Results (2000 words)


The FAO's catch statistics remain the most widely accessible and used source of information on global fisheries. Several methods have attempted to infer stock status (generally biomass relative to biomass at maximum sustainable yield, B/B_msy_) based on a fishery's catch history alone [summarized by @free2020]. However, these "catch-only" methods have serious shortcomings that limit their performance, as evidenced by the status discrepancies by region they produce (Table.\@ref(tab:status-tab)), as well as results from several studies examining the performance of such methods [@free2020; @pauly2013]. 

However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@anderson2015], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. We use a flexible assessment package ([sraplus](https://github.com/DanOvando/sraplus#sraplus)) to compare estimates of stock status around the globe using different combination of these data. We find that our perception of global stock status varies wildly depending on which data source are used, without any clear evidence of a "best" source of data to use *this needs to be much clearer*. 

## Case Study: Europe vs. Northeast Asia or East India

Don't know if this section is needed? Key messages if so (and might want to pick just one case study)

- Europe vs North East Asia
  - sort of a more positive story: adding in the Rousseau effort data at least produces a more plausible split between Europe and Northeast Asia
  - Nicely illustrates bi-modal heuristics behavior. 
  - Also lends to shortcomings though: assuming one effort series for all stocks clearly produces some weird results, e.g. all those basically unfished EU stocks

```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019. Each point is a stock, point size is a function of stock size."}
continent_kobe_plot
```


- East India
  - Local data can help: heuristic alone bi-modal and optimistic, SAR alone really pessimistic, CPUE + SAR a bit more nuanced?
  - A little bit a of a conflict here, since paper so far has been about global datasets, and here we're introducing a more local study (trawl survey abundance from eastern bay of Bengal)

```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio. Each point is a stock."}
ei_kobe_plot
```


## Conflicting Estimates of Global Status

This is the main plot of the paper Percent of stocks classified as overfished depending on what data you use. CPUE data are at the continental scale (e.g. use North America Rousseau effort for USA, Mexico, Canada). 

key message: Different data sources give you different estimates of stock status, and not in a way that makes it clear that any one is inherently superior to another


A few key questions here


- Should we use a more recent FAO report? Don't remember why we chose the 2011 report for the original report

- Currently this runs sraplus on all stocks in the region. I think this is informative since it says, suppose we only used sraplus, what would we think, and can compare that to RAM etc. in our heads (or could add a map panel). An alternative though would be to stitch together status, i.e. use RAM where there is RAM, use sraplus otherwise. That would be a better "estimate" of the actual stock status in an area, but would make the contrast between the methods less clear

- This used the continent level Rousseau data - can we get the finer scale version for this paper? Would make things a lot clearer I think 

*will add in RAM panel once I match up RAM stocks to FAO region*

```{r po-map, fig.width=8, fig.height=8, fig.cap="Estimated percent of fisheries 'overfished' (B/B_msy_ less than 0.8) using different dat sources. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
po_map_plot
```


## Value of Information

Provide some guidance on where to prioritize data collection. e.g. how long of a CPUE series is needed, expansion of f based priors / value of f, advocate for merging together of the "Catch only" world with the "length-based world". E.g. we talk about catch being the only widely available statistic, but in many data-limited communities all they have are lengths. A global effort to create a centralized database of length distributions could allow for length-based priors on spr/f that could be paired with catches?


# Discussion (500 words)

*these are just rough outline paragraphs*

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods, our understanding of many of the world's fisheries remains poor. We show that while new sources of widely available data are becoming increasingly available, none yet is capable of providing a clear improvement in our estimates of global stock status. 

Our estimate of global stock status varies wildly depending on which source of data we use. Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status? Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:po-map)). Globally available effort data from Rousseau allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region. These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability. The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations. Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks. 

We have stated that improving global estimates of stock status depends on improved data, not improved models *per say*. This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the biomass dynamics types used here, are not of use. At the tactical level, such improvements can make meaningful changes in management outcomes. However, at the global strategic levels, the improvements to be made from model development are marginal compared to the fundamental uncertainties stemming from currently available data. 

What does the future of global fishery assessment look like? Investment must be made in improving data collection at the global scale. Our analysis suggests several possible paths for this process, in particular creation of a global repository or length-composition data, and expansion of FMI surveys and SAR data. Global Fishing Watch provides another avenue, but requires further research on the best ways of translating raw effort data into indices of stock status. Interpreting these data responsibly will require support from regional experts, for example in the form of expert panels charged with determine the most representative broadly available data for different portions of the oceans. 

We have accumulated lots of evidence for status of global fisheries, each of which has strengths and weaknesses.These were important foundations. But, if we want to move forwards, the answer is better data. 

# Methods (3000 words)

## `sraplus`

Description of the package itself focused on key differences to other packages 

## Swept Area Ratio Priors

Basic description and out-of-sample predictive power

## Fisheries Management Index Priors

Basic description and out-of-sample predictive power

## Catch-Only Priors?

At the moment only using Heuristics. I've been exploring an update to the regression based catch-only models, can get up to an out of sample r^2 of about 0.3 and that's about it. Probably worth putting in just to make clear that we're making about as much use out of the catch data alone as we possible can. 

## Effort Data

Description of Rousseau data, steps taken to use it (Baranov equation, choice about tech. creep etc.)

## Value of Information Calculations

Details of value of information calculates (fitting to messed up versions of RAM)

## Estimation model

Important details of sraplus, basic equations, etc. Focus on prior-predictive tuning process to validate use of priors on terminal status. 

# References

